\vspace{-1.3cm} \textmd{\normalsize{Project description}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration

### 2016-03-27 paper/paper.tex
paper angefangen

%\pagestyle{fancy}
%\fancyhead[L]{Seminar \glqq Optimale Steuerung\grqq\\ Prof. Brokate, Prof. Kuttler} %Kopfzeile links
%\fancyhead[C]{} %zentrierte Kopfzeile
%\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
\title{\vspace{-1.3cm} \textmd{\normalsize{Project description}}\\ Parallelization of the PRM Method}
\date{}
\author{Manuel Demmeler}

@@

\section{Problem Statement and Sequential PRM Method}\label{algintro}
\newtheorem*{defi}{Definition}
\newtheorem*{alg1}{Algorithm}

The abstract problem is to find a path on a d-dimensional map, which avoids intersecting obstacles. The obstacles are  given by an indicator function
\begin{equation*}
	I: \Omega \rightarrow \{0, 1\}
\end{equation*}
on the domain \( \Omega = [ q_{min,1}, q_{max,1} ] \times ... \times  [ q_{min,d}, q_{max,d} ] \subset \R^d  \) and the goal is to find a continuous way
\begin{equation*}
	\Gamma: [0,1] \rightarrow \Omega
\end{equation*}
between a given start and endpoint, \(\Gamma(0)=q_b,\ \Gamma(1)=q_e \), 
such that
\begin{equation*}
	I(\Gamma(s))=0 \text{\ f.a.\ } sÊ\in [0,1].
\end{equation*}
The probabilistic roadmap (PRM) approach grows a graph \(G=(V,E)\) form \(q_s\) and \(q_e\) by randomly sampling nodes \(v\in V\subset \Omega\).  Two nodes \(v,w \in V\) are connected, if they are near enough each other and the linear connection in between is free from obstacles. The algorithm terminates, when there is a path from \(q_s\) to \(q_e\) on G.
\begin{defi}
For two points \(q_1\) and \(q_2\) we define the connection with stepsize \(h>0\) as
	\begin{equation*}
		[q_1,q_2]_h \defgl \{ q = \lambda q_1+(1-\lambda) q_2\ |\ \lambda \in [0,1], \norm{q-q_1} \in \N_0 h \}.
	\end{equation*}
	We say, \(q_1\) and \(q_2\) are connected (with stepsize \(h>0\)) on the map \(\Omega\), if
	\(I(q)=0\) for all \(q \in [q_1, q_2]_h\).
\end{defi}
\begin{alg1} (Probabilistic Roadmap in general)\\
	Input: \(q_s, q_e\in \Omega, h>0, d_0>0\)	\\
	Output: \(q_s=q_1, ... ,q_n=q_e\in \Omega\), 
		such that all \(q_i\) and \(q_{i+1}\) are connected with stepsize \(h\)
	\begin{itemize}
		\item Initialize graph \(G=(V,E)\) with \(V=\{q_s, q_e\}\), 
			if \(q_s\) and \(q_e\) are connected, \(E \leftarrow\{ q_s, q_e\}\)
		\item While there exists no path in G between \(q_s\) and \(q_e\) do
		\begin{itemize}
			\item Sample \(q\in \Omega\) from a probability density  \(p(\ \cdot\ , G)\), resample until \(I(q)=0\)
			\item \(V \leftarrow{q}\)
			\item For all nodes \(\tilde{q} \in V\) which are near enough (\( \norm{q- \tilde q} \le d_0 \)) and
				connected with \(q\) with stepsize h, 
				\(E \leftarrow \{q, \tilde{q}\}\)
		\end{itemize}
		\item Determine the shortest path \(q_1, ... , q_n\in V\) on \(G\) from \(q_s\) to \(q_e\).
	\end{itemize}
\end{alg1}

@@

\section{Application to Robotics}

In the project, it is planned to use this algorithm to create collisionfree trajectories of a robot arm. 
For a robot with d joint angles \(q_i\) one can calculate straightforwardly all positions \(x_j \in \R^3\) and orientations \(R_j\in \R^{3\times3}\) of the single parts of the robot, \cite{robodyn} describes the exact calculations. Therefore this direct kinematic can be interpreted as a (strongly nonlinear) function
\begin{equation*}
	f: (q_1, ... , q_d) \mapsto (x_1, R_1, ..., x_l, R_l).
\end{equation*}

Now, we have given some obstacles with fixed positions \(x_{l+1}, ..., x_m\) and orientations \(R_{l+1}, ..., R_{m}\), which the robot has to avoid during movement. It is assumed that all objects are given in a form, that for each two objects it can be determined from the positions and orientations, if there is an intersection. For example as polyhedrons.
From this the indicator function for the obstacles in the configuration space can be defined as
\begin{equation*}
	I(q)=
	\begin{cases}
		0,	& \text{if no collision for all object pairs\ } (x_j, R_j), (x_k,R_k),\ j,k\in 1, ... , m \\
		1,	& \text{for at least one collision}
	\end{cases}
\end{equation*}
Now, if we additionally introduce jointlimits and call the set of possible angles \(\Omega\), we see, that the problem of finding feasible trajectories is exactly the pathfinding problem from above.

It has to be noted that the trajectories \(q(t)\) the PRM algorithm generates are piecewise linear and therefore not differentiable. Hence, they have to be smoothed before the use on a robot. This is expectedly not part of the project.

@@

\section{Ideas for Parallelization}

@@

\bibliographystyle{alpha}
\begin{thebibliography}{999}
	\bibitem{prmlec} D. Burschka, Lecture Robot Motion Planning, source: http://robvis01.informatik.tu-muenchen.de/courses/wegtraj/index.html
	\bibitem{prm1} H. Choset, K. Lynch, S. Hutchinson, G. Kantor, W. Burgard, L. Kavraki and S. Thrun, Principles of Robot Motion: Theory, Algorithms, and Implementation, MIT Press, 2005
	\bibitem{prm2} S. M. LaValle, Planning Algorithms, Cambridge University Press, 2006
	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl fŸr Angewandte Mechanik, TU MŸnchen, 2014
\end{thebibliography}

>>>>

%\pagestyle{fancy}
%\fancyhead[L]{Seminar \glqq Optimale Steuerung\grqq\\ Prof. Brokate, Prof. Kuttler} %Kopfzeile links
%\fancyhead[C]{} %zentrierte Kopfzeile
%\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
\title{\vspace{-1.3cm} \textmd{\normalsize{Project description}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
\date{}
\author{Manuel Demmeler}

@@

\section{Introduction}
\subsection{Problem Statement and the Sequential PRM Method}\label{algintro}
\newtheorem*{defi}{Definition}
\newtheorem*{alg1}{Algorithm}

The abstract problem is to find a path on a d-dimensional map, which avoids intersecting obstacles. The obstacles are  given by an indicator function
\begin{equation*}
	I: \Omega \rightarrow \{0, 1\}
\end{equation*}
on the domain \( \Omega = [ q_{min,1}, q_{max,1} ] \times ... \times  [ q_{min,d}, q_{max,d} ] \subset \R^d  \) and the goal is to find a continuous way
\begin{equation*}
	\Gamma: [0,1] \rightarrow \Omega
\end{equation*}
between a given start and endpoint, \(\Gamma(0)=q_b,\ \Gamma(1)=q_e \), 
such that
\begin{equation*}
	I(\Gamma(s))=0 \text{\ f.a.\ } s \in [0,1].
\end{equation*}
The probabilistic roadmap (PRM) approach grows a graph \(G=(V,E)\) form \(q_s\) and \(q_e\) by randomly sampling nodes \(v\in V\subset \Omega\).  Two nodes \(v,w \in V\) are connected, if they are near enough each other and the linear connection in between is free from obstacles. The algorithm terminates, when there is a path from \(q_s\) to \(q_e\) on G.
\begin{defi}
For two points \(q_1\) and \(q_2\) we define the connection with stepsize \(h>0\) as
	\begin{equation*}
		[q_1,q_2]_h \defgl \{ q = \lambda q_1+(1-\lambda) q_2\ |\ \lambda \in [0,1], \norm{q-q_1} \in \N_0 h \}.
	\end{equation*}
	We say, \(q_1\) and \(q_2\) are connected (with stepsize \(h>0\)) on the map \(\Omega\), if
	\(I(q)=0\) for all \(q \in [q_1, q_2]_h\).
\end{defi}
\begin{alg1} (Probabilistic Roadmap in general)\\
	Input: \(q_s, q_e\in \Omega, h>0, d_0>0\)	\\
	Output: \(q_s=q_1, ... ,q_n=q_e\in \Omega\), 
		such that all \(q_i\) and \(q_{i+1}\) are connected with stepsize \(h\)
	\begin{itemize}
		\item Initialize graph \(G=(V,E)\) with \(V=\{q_s, q_e\}\), 
			if \(q_s\) and \(q_e\) are connected, \(E \leftarrow\{ q_s, q_e\}\)
		\item While there exists no path in G between \(q_s\) and \(q_e\) do
		\begin{itemize}
			\item Sample \(q\in \Omega\) from a probability density  \(p(\ \cdot\ , G)\), resample until \(I(q)=0\)
			\item \(V \leftarrow{q}\)
			\item For all nodes \(\tilde{q} \in V\) which are near enough (\( \norm{q- \tilde q} \le d_0 \)) and
				connected with \(q\) with stepsize h, 
				\(E \leftarrow \{q, \tilde{q}\}\)
		\end{itemize}
		\item Determine the shortest path \(q_1, ... , q_n\in V\) on \(G\) from \(q_s\) to \(q_e\).
	\end{itemize}
\end{alg1}

@@

\subsection{Application to Robotics}

@@

\subsection{Project Overview}

@@

\section{Structure of the PRM Solver}

here overview over the components of the programm components:

PRMSolver 1 <-> Configspace object 1 <-> GPU 1

.

.

.

PRMSolver n <-> Configspace object 2 <-> GPU 1

- description of the different versions:

- definitions as pseudocode

version 1:

	- each process generates own nodes -> exchange them, synchronous gpu kernel

version 2:

	- every process generates all nodes (same by exchange of random seed)

	- indicator function splitted over processes, synchronous

version 3:

	- same as version 2 with some pipeline optimizations

version 4:

	- same as version 3 with asynchronous gpu

version 5:

	- each process has 2 workers which are generating their nodes an are using 1 gpu

	- one worker uses the kernel time  of the other to do its work

	- should give use the gpu computation time efficiently

\section{Implementation of an Indicator Function for Robot Arms}

In the project, the probabilistic roadmap algorithm is applied  to create collisionfree trajectories of a robot arm.
For a robot with given joint angles \(q_i\) one can calculate straightforwardly all positions and orientations  of the single parts of the robot.
For this, coordinate frames according to the Denavit Hartenberg convention are used.
The coordinate frames of the single parts are described by transformation matrices 
$T_i \in R^{4x4}$, which transform points of the body frame into the world frame. They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot. For the exact calculations see \cite{robodyn}.

The geometry of the robot parts is defined by a set of convex polytopes, each of them belonging to one of the robots coordinate frame. Applying the corresponding $T_i$, its vertices can be transformed into the world frame. 
Environment obstacles are also defined by convex polytopes, in the world frame.

Now the movement of the robot can be described by a curve $q(t)$ in the state space. It is collisionfree, if at no time, any two polytopes of the robot intersect with each other or with the environment after transformed into the world frame.
To apply the probabilistic roadmap algorithm, we express this by the indicator function

\begin{equation*}
	I(q)=
	\begin{cases}
		0,	& \text{ for no collision for all given pairs of polytopes in wold frame } \\
		1,	& \text{ if at least one collision occurs }
	\end{cases}
\end{equation*}

It has to be noted that the trajectories \(q(t)\) the PRM algorithm generates are piecewise linear and therefore not differentiable. Hence, they have to be smoothed before the use on a robot. This is not part of the project. The task is only to generate feasible trajectories to be passed to an optimizer.

here some overview

- state the principle task: movement of an robot arm through environment with obstacles

- define the resulting state space: $q=(q1,...,qn) \in R^ndof$

- state how the indicator function has to look like in general

- describe the robot:

	- consisting of convex polytopes

 	- which are located in coordinate frames

	- coordinate frames described by transformation matrices

	- transformation matrices given by state q

- collision tested by the ... algorithm

\subsection{Direct Kinematics}

- coordinate frames

- DH-Parameters

\subsection{2D and 3D Geometry Library}

- vector types: float4, float2

- matrix type: trafo4=float4[4]

- float precision is enough for this application

- everything inline

- hybrid device host functions

- += etc. overloaded, + etc. to make no unnecessary temps

\subsection{Polytopes}

- describe polytope4 structure:

	- vertices stored as float4 array

	- edges stored in crs format

- whole list of polytopes stored compactly in polytope4data on gpu

- maybe describe interface polytope <-> polytope4data <-> polytope4

\subsection{Chung-Wang Collision Algorithm}

- input: polytopes P, Q and trafonsformations Tp, Tq

- output: intersection or not

- describe very shortly how the algorithm works

- implemented as simple device function

\subsection{Robot Kernel}

- the indicator function gets:

	- polytopes in coordinate frames

	- data, which pairs must be tested

	- q

  and computes if for any pair $(i,j)$ the polytopes $P_i, P_j$ collide in the world frame

- implements the indicator function for multiple requests parallely:

- input: lists qstart, qend

- output: res

$res[i] = 0$ if indicator(q) == 0 for all q in $[ qstart[i], qend[i] ]_deltaq$

	!= 0 else

program:

host:

- computes distances between qstart, qend and resulting number of intermediate points/ threads

- qstart, qend, numbers of threads sendt to gpu

gpu kernel:

- each gpu thread computes one intermediate point and its indicator function

- indicator function computation:

	- iterate over all polytope pairs and run the collision algorithm for each

- reduce the indicator functions to res

- send back res to host

--> maybe do this in pseudo code

asynchronous version:

- indicator call: everything untill kernel call

- wait function: res memcopy incl. waiting for kernel

\bibliographystyle{alpha}
\begin{thebibliography}{999}
	\bibitem{prmlec} D. Burschka, Lecture Robot Motion Planning, source: http://robvis01.informatik.tu-muenchen.de/courses/wegtraj/index.html
	\bibitem{prm1} H. Choset, K. Lynch, S. Hutchinson, G. Kantor, W. Burgard, L. Kavraki and S. Thrun, Principles of Robot Motion: Theory, Algorithms, and Implementation, MIT Press, 2005
	\bibitem{prm2} S. M. LaValle, Planning Algorithms, Cambridge University Press, 2006
	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl für Angewandte Mechanik, TU München, 2014
\end{thebibliography}
